---
title: "read the `stars.txt` file"
output: html_notebook
---


```{r setup}
library(tidyverse)

library(datapasta)
```

This is a nasty case...the source data is from here.

[Star Charts and Constellations](http://nifty.stanford.edu/2009/reid-starmap/starmap.html)


The description of the file is here:

* The first three fields are the x, y and z coordinates for the star. We will ignore the z coordinate, and use only the x and y coordinates. Each axis in the coordinate system goes from -1 to +1, and the centre point is 0,0. (See the figure below.)

* The fourth field is the Henry Draper number, which is simply a unique identifier for the star.

* The fifth field is the magnitude (or brightness) of the star.

* The sixth field is Harvard Revised number, another identifier.

* The seventh field exists only for a small number of stars and is a semicolon-separated list of names for a star. A star may have several names.

The issue is with the last column.


I started with {readr} but this doesn't quite work; the issue originates with the last column.


A basic read drops the last column entirely:

```{r}

read_delim("stars_10.txt", delim = " ") 

```

Specifying the last column as a character doesn't help, since it pre-parses to the 6 columns, and then bonks out on the 7 and 9 column structure (note that row 6 last column is "CAPH; CAS BETA" where the semi-colon is the separator and the 2nd name has a space).

```{r}

stars <- read_delim("stars_10.txt", delim = " ", 
                      col_types = cols(
                        x = col_double(),
                        y = col_double(),
                        z = col_double(),
                        HDnum = col_double(),
                        mag = col_double(),
                        HRnum = col_double(),
                        name = col_character()
                        )
)


```


So I tried defining the names, which got me closer. By using `col_names` I got to a place where it was reading in at least the first part of the last column.


```{r}

read_delim("stars_10.txt", delim = " ", col_names = FALSE)


read_delim("stars_10.txt", delim = " ", 
 col_names = c("x", "y", "z", "HDnum", "mag", "HRnum", "name"))
   

```


So I extended the number of "name_" variables:


```{r}


read_delim("stars_10.txt", delim = " ", 
 col_names = c("x", "y", "z", "HDnum", "mag", "HRnum", "name_a", "name_b", "name_c"))
   

```
 
This works here, but a quick scan of the main file reveals cases where there are three names (i.e. two semi-colons), two of which are doubles like "CAS BETA". I didn't explore the entire universe...

I thought about using `readr::read_lines` but I don't see an easy way to parse this.

```{r}

readr::read_lines("stars_10.txt")

```



I also tried the {readtext} package but that was a dead-end as well.

* https://readtext.quanteda.io/



```{r}

readtext::readtext("stars_10.txt")

```


Based on a stackoverflow question the `read.table()` from {data.table} seemed like a possibility, but it then wraps the bonus text as V1 etc...

```{r}

read.table("stars.txt", header = FALSE, sep = " ",
           col.names = paste0("V", seq_len(7)), fill = TRUE)

```


This might require a regex solution, or maybe creating an index variable (based on the row names), splitting the file into cols 1-6 as neatly parsed columns, the names (where they exist) as a second, using regex to fix the latter, then joining them back together.


